---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: '2024-02-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hierarchical Clustering

read case

```{r}
#remember to change the path
rm(list=ls())
library(rvest)
library(tidyverse)
library(readxl)
setwd("D:\\Thesis\\sample\\dp-s88-phen")
set1 <- read_excel("DepressionGenesNetworks_AllSCIDS_distribution.xlsx", sheet = "DATA", col_names = T)
set2 <- read_excel("DepressionGenesNetworks_case_control_phenotype_distribution.xlsx", sheet = "Data", col_names = T)
set3 <- read_excel("DepressionGenesNetworks_ScreenData_distribution.xlsx", sheet = "NIH_Depression_Hopkins_18NOV10", col_names = T)
colnames(set2)[3] <- "cell_id"
#remove replicate variables in  set1 and set2
set2_reduced <- set2[,!colnames(set2) %in% colnames(set1)[colnames(set1) != "cell_id"]]
data <- merge(set1,set2_reduced,by="cell_id")
#remove replicate variables in  set2 and set3
set3_reduced <- set3[,!colnames(set3) %in% colnames(data)[colnames(data) != "cell_id"]]
data <- merge(data,set3_reduced,by="cell_id")

data <- data[data$cell_id!="LD0053",]

control <- data[data$case_control_status==1,]
case <- data[data$case_control_status==2,]
```

save gender

```{r}
setwd("D:\\Thesis\\WORKSPACE2024\\Hierarchical Clustering\\data")
gender <- case$gender
save(gender,file = "gender.RData")
```

read Encoder

```{r warning=FALSE}
setwd("D:\\Thesis\\WORKSPACE2024\\variables exploration\\data")
library(umap)
library(FactoMineR)
library(factoextra)
rm(list=ls())
encode <- read.table("case_encode.txt",header=T,sep = "\t",row.names = NULL)
round_encode <- as.data.frame(lapply(encode, function(x) round(x,digits = 4)))
```

Scaling

```{r}
round_encode <- as.data.frame(scale(round_encode))
```

read gender

```{r}
setwd("D:\\Thesis\\WORKSPACE2024\\Hierarchical Clustering\\data")
load("gender.RData")
```

divide into female dataset and male datset

```{r}
female_encode <- round_encode[gender=="F",]
male_encode <- round_encode[gender=="M",]
```

### Female

apply euclidean distance method to construct distance matrix

```{r}
dist_mat_f <- dist(female_encode, method = 'euclidean')
```

use Complete-linkage linkage method

```{r}
hclust_avg_f <- hclust(dist_mat_f, method = 'complete')
plot(hclust_avg_f,labels=FALSE)
```

explore the best clustering number

```{r}
fviz_nbclust(female_encode, hcut, method = "wss")+
  geom_vline(xintercept = 2, linetype = 2)
```

divide into 2 clusters

```{r}
plot(hclust_avg_f,labels=FALSE)
cluster_f <- rect.hclust(hclust_avg_f , k = 2, border = 2:6)
```

### Male

apply euclidean distance method to construct distance matrix

```{r}
dist_mat_m <- dist(male_encode, method = 'euclidean')
```

use Complete-linkage linkage method

```{r}
hclust_avg_m <- hclust(dist_mat_m, method = 'complete')
plot(hclust_avg_m,labels=FALSE)
```

explore the best clustering number

```{r}
fviz_nbclust(male_encode, hcut, method = "wss")+
  geom_vline(xintercept = 2, linetype = 2)
```

divide into 2 clusters

```{r}
plot(hclust_avg_m,labels=FALSE)
cluster_m <- rect.hclust(hclust_avg_m , k = 3, border = 2:6)
```

### Save Cluster

cluster index

```{r}
cluster1_index <- as.numeric(names(cluster_f[[1]]))
cluster2_index <- as.numeric(names(cluster_f[[2]]))
cluster3_index <- as.numeric(names(cluster_m[[1]]))
cluster4_index <- as.numeric(names(cluster_m[[2]]))
cluster5_index <- as.numeric(names(cluster_m[[3]]))
```

cluster assignment

```{r}
cluster_assignment <- c()
cluster_assignment[cluster1_index] <- rep(1,238)
cluster_assignment[cluster2_index] <- rep(2,121)
cluster_assignment[cluster3_index] <- rep(3,1)
cluster_assignment[cluster4_index] <- rep(4,71)
cluster_assignment[cluster5_index] <- rep(5,31)
```

save cluster assignment

```{r}
setwd("D:\\Thesis\\WORKSPACE2024\\Hierarchical Clustering\\data")
save(cluster_assignment,file = "cluster_gender.RData")
```
